Fast Sorting on Modern Computers

### 2005-08-14 docs/tech_report/radix.tex
renamed cases of 2[ \-]level adaptive to two-level
fixed cache misses slope vs amortization thing

Again we see that the bimodal branch predictor performs better than the two
level adaptive. In this case it is contrary to expectations, however, as
there are no comparative branches. However, there are so few misses that these
results do not show anything; in the worst case there are slightly over 200
misses in total.

@@

LaMarca's radixsort results are quite different to those presented here. His
instruction count graph has a logarithmic shape: it begins very high and sharply
falls before flattening out. By contrast, the graph here is almost completely
flat. He shows a similar result in both cache miss graphs: a sharp dip, followed
by a nearly flat-line. Also, in both cases, once the array no longer fits in the
cache, the number of misses jumps to around 2.5, where they stay.

However, LaMarca's conclusion - that radixsort is not as fast as quicksort - can
be easily seen from the base radixsort results. Once an radixsort optimised for
the cache is presented, these no longer hold, and the lack of branch
misprediction results in a new fastest sort.

>>>>

Again we see that the bimodal branch predictor performs better than the
two-level adaptive. In this case it is contrary to expectations, however, as
there are no comparative branches. However, there are so few misses that these
results do not show anything; in the worst case there are slightly over 200
misses in total.

@@

LaMarca's radixsort results are quite different to those presented here. His
instruction count graph has a logarithmic shape: it begins very high and sharply
falls before flattening out. By contrast, the graph here is almost completely
flat. He shows a similar result in both cache miss graphs: a sharp dip, followed
by a nearly flat-line. This may be due to amortization, in which case the trend
would not be expected in our graphs due to subtraction of the time taken to fill
the array.

Also, in both cases in LaMarca's graphs, once the array no longer fits in the
cache, the number of misses jumps to around 3.25 (which would be equivalent to
1.625 misses per key in our results), where they stay. Our results
disagree, showing approximately 2.5 misses per key for base radixsort.

However, LaMarca's conclusion - that radixsort is not as fast as quicksort - can
be easily explained from the base radixsort results. Once an radixsort optimised
for the cache is presented, these no longer hold, and the lack of branch
misprediction results in a new fastest sort.
